---
title: "BDA - Project"
author: "Atul Kumar"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 1
---

# Introduction 
Problem of Ranking appear when we want to know the underlying order of given data. For example given a list of movie on netflix and each user preference we want to know which movie comes in top 10. The chalenge here is that we don't have preference form each user for each movie. We know a small movie preferences for each user. 

The problem is to figure out a universal order of movie ranking given a set of preferences provided by the user.


# Dataset
For this experiment I used data from https://www.preflib.org/data/election/netflix/
I used .soc files 1 - 100 which all have 3 candidate elections.
All .soc files are merged into one dataframe and movie names are mapped to index.
The processing files are found here https://github.com/atulkum/BDA_project/blob/main/process_data.py

Overall there are 171 movies (indicated by N_options in the code). Also we have around 600 rows in the data each has a weight and 3 elections order of the movies given by the user. 

```{r, eval = FALSE}
1) movie name to id

id movie_name
-------------
1	Shrek (Full-screen)		
2	The X-Files: Season 2		
3	The Punisher		
4	Spy Game		
5	Lethal Weapon 4		
6	Glengarry Glen Ross

2) User Ranking data

count rank1 rank2 rank3
-----------------------
263	2	1	3
249	1	2	3
78	1	3	2
46	2	3	1
17	3	1	2
11	3	2	1
```

# Model 
For this problem I used "Exploded Logit" https://en.wikipedia.org/wiki/Discrete_choice#exploded_logit 

The stan code is taken from the this link and modified for weighted ranking data and used in netflix dataset. https://bnicenboim.github.io/2021/03/21/a-simple-way-to-model-rankings-with-stan/

# Prior 
The uniform Dirichlet distribution is used as prior, which indicate we are not certain that how sparse the probabilities are. The ranking score varibale theta is declared as a simplex so that all probability should sum to 1.

# Stan model code

```{r}
cat(readLines('./weighted_exploded.stan'), sep = '\n')
```


# inference from models

Following is the inference code for the model. 4 chains are used and each chain run for 2000 iteration of which 1000 is thrown awaya as warmup.

```{r, eval = FALSE}
library(dplyr)
library(purrr) 
library(ggplot2) 
library(posterior)
library(cmdstanr) 
library(bayesplot) 

ldata <- list(res = as.matrix(netflix_data %>% select(starts_with("rank"))),
              weights = netflix_data$count, 
              N_options = nrow(movie_names), 
              N_ranked = ncol(netflix_data)-1, 
              N_ranking = nrow(netflix_data)) 

m_expl <- cmdstan_model("./weighted_exploded.stan")

f_exploded <- m_expl$sample(
  data = ldata,
  seed = 42,
  parallel_chains = 4
)
saveRDS(f_exploded, file = "f_exploded.RDS")
```

# Result

Following code the movie names are ranked by mean score of the draws and top 10 movies come out are

111	Lost in Translation		
113	Men in Black II		
115	Jurassic Park		
73	Swordfish		
94	Monster's Ball		
95	Seven		
39	The Silence of the Lambs		
1	  Shrek (Full-screen)		
59	I		
28	Say Anything

```{r, eval = FALSE}
f_exploded <- readRDS("f_exploded.RDS")
movie_rankings = data.frame(
  by_mean = movie_names[order(theta_summary$mean, decreasing = TRUE)[1:10],"movie_name"],
  by_median = movie_names[order(theta_summary$median, decreasing = TRUE)[1:10],"movie_name"],
  by_mad = movie_names[order(theta_summary$mad, decreasing = TRUE)[1:10],"movie_name"],
  by_p95 = movie_names[order(theta_summary$q95, decreasing = TRUE)[1:10],"movie_name"]
)
movie_rankings
```


# Simulation - Based calibration
To find the issues in the model I used "Simulation - Based calibration"
The code used for this is given below.

Somehow the chi squre test for uniformality is giving 0 as p-value. I still need to investigate the model.

Reference https://mc-stan.org/docs/2_25/stan-users-guide/testing-uniformity.html

```{r}
cat(readLines('./weighted_exploded_scb.stan'), sep = '\n')
```

## Simulation

```{r, eval = FALSE}
df_all = NULL
M = 999
for (i in 1:M) { 
  data_sim <- list(
                N_options = nrow(movie_names), 
                N_ranked = ncol(netflix_data)-1, 
                N_ranking = 10) 
  
  
  m_expl_sim <- cmdstan_model("./weighted_exploded_scb.stan")
  
  f_exploded_sim <- m_expl_sim$sample(
    data = data_sim,
    parallel_chains = 4,
    refresh = 0
  )
  rank_stat_df <- as_draws_df(f_exploded_sim$draws("ranks_")) %>% select(starts_with("ranks_"))
  if (i == 1){
    df_all <- colSums(rank_stat_df)
  } else {
    df_all <- rbind(df_all, i=colSums(rank_stat_df))
  }
}
rownames(df_all)<-NULL
```

## binning
```{r, eval = FALSE}
J <- 50
n_choices <- 171
M_draw <- 4000
b <- matrix(0L, nrow = J, ncol = n_choices)
r <- matrix(as.numeric(unlist(df_all)),nrow=nrow(df_all)) #999 171
for (k in 1:n_choices){
  for (m in 1:M){
    idx <- 1 + floor(r[m, k] * J / (M_draw + 1))
    b[idx, k] <- 1 + b[idx, k]
  }
}
```

## Chi-suqre test for uniformaliy

```{r, eval = FALSE}
for (k in 1:n_choices){
  ej = (M_draw + 1)/J
  chi_square_value = floor(sum((b[, 1] - ej)^2/ej))
  print(pchisq(chi_square_value, df=J-1, lower.tail=FALSE))
}
```

